Practical Optional Types for Clojure

### 2015-02-19 formal-model.tex
intro section 3

@@

                &\mbox{Constants}\\
  \s{}, \t{}    &::=& \Top 
                      \alt {\Unionsplice {\overrightarrow{\t{}}}}
                      \alt
                      {\ArrowOne {\x{}} {\t{}}
                                   {\t{}}
                                   {\filterset {\prop{}} {\prop{}}}
                                   {\object{}}}
                      \alt {\Value \singletonmeta{}} 
                      \alt \trdiff{\class{}}
                &\mbox{Types} \\
  \singletonallsyntax{}
                \\ \\
  \occurrencetypingsyntax{}
  %\\
  %\classliteralallsyntax{}
\end{altgrammar}
$$
\caption{Syntax of Terms, Types, Propositions and Objects}
\end{figure*}

@@

    {\TSubsume}
  \end{mathpar}
  \caption{Standard typing rules}
\end{figure*}

@@

    \BDelta{}
  \end{mathpar}
  \caption{Semantics}
\end{figure*}

\begin{figure*}
  \footnotesize
\begin{mathpar}
\standardsubtyping{}
\end{mathpar}
\caption{Subtyping rules}
\end{figure*}

>>>>

Syntax is given in figure~\ref{figure:termsyntax}. Expressions include variables, values,
application, abstractions, conditionals, do (sequencing) expressions and let expressions.
We model throwable errors without a catching mechanism. All binding forms introduce fresh
variables.

Values include booleans, \nil{}, class literals, keywords, constants and closures.
\false{} and \nil{} are the only logically false values.

Types include the top type, untagged unions, functions, singleton types
and class instances. We abbreviate \EmptyUnion{} as \Bot{}.
Function types contain \emph{latent} propositions and object, which, along with the return type,
may refer to the parameter identifier. They are latent because they are instantiated with the
actual object of the argument in applications before they are used in the proposition environment.

Propositions are identical to~\cite{TF10}; positive and negative type propositions, implications,
conjunctions, disjuctions, trivial and impossible propositions.
Type propositions contain a non-empty object which it claims type information about.

Proposition environments are sequences of propositions in conjunctive normal form.

Typing rules are in figure~\ref{figure:typingrules}.
T-Error neither returns logically true or false as it does not return normally.
T-Do uses the propositions learnt in the first expression to type check the second.
Other than compensating for the extra logically false value \nil{}, the rest are standard.

Figure~\ref{figure:standardopsem} contains the operational semantics, which are standard
except for \nil{} being false.

Subtyping (figure~\ref{figure:subtyping}) says singleton types are instances of their respective classes
and \Object is transitively a supertype to everything except {\Nil{}},
so \Top{} = {\Union{\Nil}{\Object}}.

@@

                &\mbox{Constants}\\
  \s{}, \t{}    &::=& \Top 
                      \alt {\Unionsplice {\overrightarrow{\t{}}}}
                      \alt
                      {\ArrowOne {\x{}} {\t{}}
                                   {\t{}}
                                   {\filterset {\prop{}} {\prop{}}}
                                   {\object{}}}
                      \alt {\Value \singletonmeta{}} 
                      \alt \trdiff{\class{}}
                &\mbox{Types} \\
  \singletonallsyntax{}
                \\ \\
  \occurrencetypingsyntax{}
  %\\
  %\classliteralallsyntax{}
\end{altgrammar}
$$
\caption{Syntax of Terms, Types, Propositions and Objects}
\label{figure:termsyntax}
\end{figure*}

@@

    {\TSubsume}
  \end{mathpar}
  \caption{Typing rules}
  \label{figure:typingrules}
\end{figure*}

@@

    \BDelta{}
  \end{mathpar}
  \caption{Operational Semantics}
  \label{figure:standardopsem}
\end{figure*}

\begin{figure*}
  \footnotesize
\begin{mathpar}
\standardsubtyping{}
\end{mathpar}
\caption{Subtyping rules}
\label{figure:subtyping}
\end{figure*}
